---
title: "Super rollin' along with Conditional Means"
author: "Garcia"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls()) #this clears the global environment - if you don't want to clear your GE, then comment this code out (or delete it altogether)

library(tidyverse)
library(rvest)
library(tigris)
library(lubridate)
library(readxl)
library(yardstick)

# Get File from Github
if(file.exists("SB_R.xlsx")==FALSE){
  download.file("https://github.com/rafael-a-garcia/LLO8200_updated/raw/main/modules/SB_R.xlsx", destfile="SB_R.xlsx", mode="wb")
  sBowl<-read_excel("SB_R.xlsx", skip=3, col_names=TRUE)  
  }else{
  sBowl<-read_excel("SB_R.xlsx", skip=3, col_names=TRUE)
}

# Convert Possession time into seconds, format date so R recognizes is, Pull out year
sBowl<-sBowl%>%
  mutate(Possession=ms(as.character(Possession)))%>%
  mutate(Possession=period_to_seconds(Possession))%>%
  mutate(Date=ymd(sBowl$Date))%>%
  mutate(Year=year(sBowl$Date))

```
| Name | Definition |
| ------- | --------- |
|Date | Date of Superbowl
|SuperBowl | Number of Superbowl|
|Team | Team that played|
|WonGame | Whether that team won|
|Points | How many points that team scored|
|FirstDowns | # of first downs|
|RushAttempts | # of Rushing attempts|
|RushYards | # of Yards earned through Rushing|
|RushTDs | # of Touchdowns Rushed|
|Fumbles | # of Fumbles|
|FumblesLost | # of Fumbles lost|
|PassAttempts | # of Passing attempts|
|PassCompleted | # of Completed passes|
|PassYards | # of Yards earned through Passing|
|PassTDs | # of Touchdowns Passed|
|Interceptions | # of Interceptions caused|
|TotalYards | # of Yards earned Total|
|Possession | Total time ball was in possession|
|HF | Whether or not the MVP went on to be a Hall of Famer|
|City | City game took place|
|State | State game took place|
|Flip | Coin Flip: Heads/Tails|
|WonFlip | Whether team won the flip|
|Year | Year game took place|

# Your Challenge:

Use What you know about Conditional Means to build a model predicting the winner of the game (`WonGame`). You can create any variables you need to from this dataset (ex: you want to look at the # of Fumbles Lost per Fumble, you can mutate `Fumbles Lost` divided by `Fumbles`), but remember to watch for linear dependancy (you  shouldn't calculate the `Interceptions` per `TotalYards` AND also include `TotalYards` as a predictor). Some variables are already categorical, others will need to be `ntile`d or something else. 

The group that can build the "best model" will earn 5 EC points to be awarded to the next Problem Set it can be. "Best" will be determined by the instructor and will consider: RMSE, # of predictors in the model, # of levels total. In all three of these cases: a lower number is better. Make sure to include graphs of your model's ability to predict. 



```{r}
test<-sBowl%>%
  mutate(avg = mean(WonGame, na.rm = TRUE))%>%
  ungroup()


test%>%
  rmse(WonGame, avg)

#Unconditional Mean is .50

test<-test%>%
  group_by(WonFlip)%>%
  mutate(avg_wflip = mean(WonGame, na.rm = TRUE))%>%
  ungroup()


test%>%
  rmse(WonGame, avg_wflip)

#Compare to unconditional mean: (New RMSE - Old RMSE)/Old RMSE
(0.496904	-.5)/.5

test<-test%>%
  group_by(HF)%>%
  mutate(avg_hf = mean(WonGame, na.rm = TRUE))%>%
  ungroup()


test%>%
  rmse(WonGame, avg_hf)

(0.4179698	-.5)/.5

```

**A Statistical Aside:** Strictly speaking, you wouldn't normally use the RMSE to assess the performance of a model with a binary outcome. You would usually use more traditional classification metrics like sensitivity and specificity; however, we won't be talking about those until much later. It turns out there *is* a metric for this kind of task called a Brier Score. Its calculation is functionally equivalent to the MSE when the outcome is unidimensional (as it is here). Therefore, the Brier Score is just the RMSE squared. 